\section{Conclusion}

In this review, the multi-dimensional search problem has been defined. Problems from many different areas of computer science, such as database querying, computer vision and visualisation, benefit from multi-dimensional search, meaning there is much benefit to gain from accelerating this process. A large array of index structures have been developed throughout the past forty years to solve this problem. Sequential scan was deemed an inefficient, na\"{i}ve solution to the problem and more sophisticated index structures were developed.

Each index structure has its own advantages and disadvantages because they were developed to tackle specific challenges. While some structures generally outperform others, there is often no ``best" structure. When considering which to use for a given application, it is important consider some key aspects about the data and the application it is being used in (see Section \ref{sec:structure-decision}). The answers to these questions can help guide the decision on which index structure to use.

The one-dimensional search problem is generally well solved and fast, dynamic index structures capable of storing huge amounts of data exist. For a low number of dimensions (e.g. 2 or 3), research has focused more on developing simple index structures with lightweight memory requirements. For high-dimensional data, focus appears to be shifting away from tree-based structures that recursively decompose space and towards linear or hash-based structures. This is because tree-based approaches have been shown to have limited, if no, performance gain over sequential-based approaches when dealing with large numbers of dimensions \cite{md-structures-samet}. There has been research into parallelising search, but it appears to remain a difficult problem, with little focus being given to parallelising high-dimensional search specifically.
