\section{Progress Report}

Referring to the schedule defined in Section \ref{sec:schedule}, as of the submission of this report, the following phases are complete: ``Project Definition", ``Literature Review and Data Collection" and ``Mid-project Presentation and Report". The following milestones have also been met: \textbf{Project Outline Defined}, \textbf{Literature Review Complete},  \textbf{Test Data Collected}, \textbf{Mid-project Presentation Delivered} and \textbf{Mid-project Report Finished}. 

This means that the project's aims, objectives and requirements have been defined, with a clear description on what the final deliverables are. Extensive research has been performed on the multi-dimensional search problem, existing solutions, their limitations and general challenges the field faces. All this information was written up into a stand-alone literature review document that has been integrated into Section \ref{sec:background-research} of this report. Which index structure to implement initially has been determined and the technologies to use to develop the implementation have also been decided. The development process has been determined which, as discussed in Section \ref{sec:schedule}, will involve multiple iterations of implementation, testing and evaluation. Finally, a presentation has been performed to the relevant research group, Computational Science and Engineering, outlining the project's aims, current progress and the design, implementation and evaluation approach which will be taken throughout the next two months.

See Section \ref{sec:next-steps} for what actions will be performed next, after the submission of this report.
